---
title: "Scrape 2 finished"
output: 
---

Scrape the warn notices for the state of South Dakota, from <https://dlr.sd.gov/workforce_services/businesses/warn_notices.aspx>.

This scrape, while similar to the MD WARN, presents the problem of scraping links from the table. Links are stored as attributes inside of <a> tags, so you'll use the html_attr() function to pull them out. To start, though, the code is similar to MD WARN: 

```{r}
library(tidyverse)
library(rvest)
```

This section of the code is nearly identical:
```{r}
url <- "https://dlr.sd.gov/workforce_services/businesses/warn_notices.aspx"

html <- read_html(url)

html_table <- html |> html_elements("table")

final_table <- 
  html |> html_element("table") |> 
  html_table(header=T)
```

At this point if you use html_table() you will wipe out all attributes. You need to extract them first.
The links live inside an <a> tag within the first <td> tag for each row: 
`<a href="/workforce_services/businesses/warn_notices/target-corp-sioux-falls.pdf" title="Target Corporation WARN Notice" target="_blank">Target Corporation</a>`. It's the "href" attribute that holds the actual link. Use html_elements() to find the <a> tags and html_attr() to pull the link associated with "href":

```{r}
# test it out:
html_table |>
  html_elements("a") |> 
  html_attr("href")

# pull out the links into their own column:
links <- html %>% 
  html_elements("table a") %>% 
  html_attr("href")

# because there aren't links for every row, only the first 48, we need to fill out the data with NAs using replicate():
links_col <- append(links, replicate(20, NA)) %>% as_tibble()

# bind this column of links to the original table: 
final_table <- bind_cols(final_table, links_col)
```

